---
title: "2017-10-16 HW4_Q1"
author: "Ethan M. Lotz"
date: "October 16, 2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("http://www.bioconductor.org/biocLite.R")
library(plyr)
library(dplyr)
library(biomaRt)
library(annotables)
library(data.table)
library(Biobase)
options(stringsAsFactors = FALSE)
```

## Question 1
Create expression matrix
```{r}
Q1 <- read.table("63_immune_cells/data/counts.txt.gz", header=TRUE)
colnames(Q1) <- sub(".bam", "", colnames(Q1))
```

Note the first 6 columns are gene annotation columns. Extract them into a separate variable - this
is a part of your feature annotation data
```{r}
Q1Annotation <- subset(Q1, select=c(1:6))
```

Extract the other columns into a separate variable - this is your expression matrix
```{r}
ExpressionMatrix <- subset(Q1, select = c(7:69))
```

Add row names (probe IDs) to the expression matrix, extracted from column “Geneid”.
```{r}
rownames(ExpressionMatrix) <- Q1$Geneid
```

What type of probe ID is it?
```{r}
print("Ensemble")
```

Create feature annotation matrix
```{r}
Q1Annotation <- Q1Annotation  %>% 
    dplyr::left_join(grch37, by = c("Geneid" = "ensgene")) %>% 
    dplyr::select(Geneid, chr, Start, End, Strand, Length, symbol, description)
```

How many probe IDs did not have mapping?
```{r}
sum(is.na(Q1Annotation$symbol))
```

How many probe IDs have description but no gene symbol?
```{r}
sum(is.na(Q1Annotation$symbol) & Q1Annotation$description!="", na.rm=TRUE)
```

Create sample annotation matrix
```{r}
SampleAnnotation <- read.table("63_immune_cells/data/E-MTAB-2319.sdrf.txt", sep="\t", header=TRUE, as.is=TRUE)
SampleAnnotationMatrix <- data.frame(row.names = colnames(ExpressionMatrix), SampleAnnotation[match(colnames(ExpressionMatrix), SampleAnnotation$Comment.ENA_RUN.), ])
```

How many different cell types are there?
```{r}
data.table(SampleAnnotationMatrix)[, uniqueN(Factor.Value.cell.type.)]
```

How many replicates per cell type?
```{r}
ddply(SampleAnnotationMatrix,.(Factor.Value.cell.type.),nrow)
```

Create a minimal ExpressionSet using expression (assayData) and sample annotation (phenoData)
matrices. Print the resulting ExpressionSet
```{r}
phenoChar1 <- new("AnnotatedDataFrame", data = SampleAnnotationMatrix)
MinimalSet <- ExpressionSet(assayData = as.matrix(ExpressionMatrix), phenoData = phenoChar1)
MinimalSet
```


